---
title: "Metabarcoding Singapore <br> ASV from Dada2"
author: "Daniel Vaulot, Adriana Lopes dos Santos"
date: '`r format(Sys.time(), "%d %m %Y")`'
header-includes:
   - \usepackage{color, fancyvrb}
output:
  pdf_document: 
    toc: yes
    toc_depth: 3
    number_sections : yes
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    number_sections : yes    
---



# Initialize. 

This file defines all the necessary libraries and variables

```{r init, eval=TRUE, message=FALSE, warning=FALSE}
  source('Metabarcoding Singapore_init.R', echo=FALSE)
```

# Aim

* Assign and analyze eukaryotes for Singapore metabarcoding data (ASV assigned with dada2 as implemented on Mothur).  
* Do some analyzes with the prokaryotes too...


# Read the data
## File names

```{r}
  full_path_data <- function(file_name) { str_c("../qiime/2018-09-06_dada2/", file_name)}

  taxo_file <- full_path_data("taxonomy.tsv")
  otu_file <- full_path_data("feature-table_unrarefied.tsv")
  sequence_file <- full_path_data("ref_sequences_unrarefied.fasta")
  
  metadata_file <- "../metadata/Singapore_metadata.xlsx"
  
  sequence_file_euk <- full_path_data("ASV_unrarefied_euk.fasta")
  dada2_taxo_file_euk <- full_path_data("ASV_unrarefied_euk.dada2.taxo")
  dada2_boot_file_euk <- full_path_data("ASV_unrarefied_euk.dada2.boot")
  otu_table_final_file <- full_path_data("ASV_final.tsv")
  
  blast_file <- full_path_data("ASV_unrarefied_euk.blast.tsv")  
```


## Read the files
* The dada2 treatment has already removed the forward and reverse primers, so no need to remove them
* Work with the unrarefied data

```{r}
# Read the sample and metadata tables
  sample_table <- read_excel(metadata_file, sheet="samples", range="A1:D89") 
  metadata_table  <- read_excel(metadata_file, sheet="metadata", na=c("ND", ""))
  sample_table <- left_join(sample_table, metadata_table) %>% 
                  mutate(sample_label = str_c(strait_label,location_label,
                                              monsoon,sprintf("%03d",day_number), 
                                              sep="_"))

# Read the taxonomy table
  taxo_table <- read_tsv(taxo_file)

# Clean up the taxonomy
  taxo_table <- taxo_table %>% 
    mutate(taxo_clean = str_replace_all(Taxon, "D_[0-9]+__","")) %>% 
    separate(col=taxo_clean, into=str_c("taxo", c(1:7)), sep=";") %>% 
    rename(otu_name = `Feature ID`)
  
# Read the otu table
  otu_table <- read_tsv(otu_file, skip=1) %>%  # Jump the first line
    rename(otu_name = `#OTU ID`) %>% 
    mutate(otu_id = str_c("otu_", sprintf("%04d",row_number())))

# Read the sequences
  otu_sequences <- readAAStringSet(sequence_file)
  otu_sequences.df <- data.frame (otu_name=names(otu_sequences),sequence=as.character(otu_sequences))

# Remove the primers - Not necessary because the primers have been removed  
  # fwd_length = 20
  # rev_length = 15
  # otu_sequences.df <- otu_sequences.df %>% 
  #   separate (col=names, into=c("otu_id_qiime", "otu_rep_seq"), sep=" ") %>%
  #   mutate (sequence = str_sub(sequence, start=fwd_length+1, end = - rev_length - 1))

  otu_table <- taxo_table %>%  
    left_join(otu_table) %>% 
    left_join(otu_sequences.df) %>% 
    arrange(otu_id)
  
# Write a fasta file for blast with all taxonomy roups
 # otu_sequences <- otu_table %>% transmute(sequence=sequence, seq_name=otu_id)
 # fasta_write(otu_sequences, file_name="../qiime/otu_rep_98_all.fasta", compress = FALSE, taxo_include = FALSE)  
  
  
```

## Only keep the eukaryotes in the OTU file

```{r}
 otu_table_euk <- otu_table %>% filter(str_detect(Taxon, "Eukaryota"))

 # Write the fasta file file
 otu_sequences_euk <- otu_table_euk %>% transmute(sequence=sequence, seq_name=otu_id)
 fasta_write(otu_sequences_euk, file_name=sequence_file_euk, 
             compress = FALSE, taxo_include = FALSE)
 
```
# Assignment of eukaryotic ASVs based on PR2 database

## Use dada2 to reassign to PR2

```{r, eval=FALSE}
 dada2_assign(seq_file_name=sequence_file_euk)
 
```

## Read the PR2 assignement and merge with initial otu table

```{r}
 otu_euk_pr2 <- read_tsv(dada2_taxo_file_euk)
 otu_euk_pr2_boot <- read_tsv(dada2_boot_file_euk) %>%
   rename_all(funs(str_c(.,"_boot"))) %>% 
   rename(seq_name = seq_name_boot)
 otu_euk_pr2 <- left_join(otu_euk_pr2, otu_euk_pr2_boot) %>% 
   rename(otu_id= seq_name)
 otu_table_final <- left_join(otu_table, otu_euk_pr2) %>% 
   select(otu_id, otu_name,taxo1:taxo7, Taxon, kingdom:species_boot,matches("EC|PR|RM|SBW|STJ"),sequence)
          
 write_tsv(otu_table_final, otu_table_final_file, na = "")
```
# Process BLAST file

BLAST is performed on Roscoff ABIMS server

```{r, eval=FALSE}
blast_18S_reformat(blast_file)
```


# Phyloseq analysis

## Create phyloseq files for euk after filtering the data

Filter the euk data to remove the low bootstraps values (threshold : bootstrap > 90% at the supergroup level) and create a phyloseq file

Note the bootstrap threshold had to be higher for 98% compared to 97% (90% vs 65%)
```{r warning=FALSE}
  otu_table_euk_final <- otu_table_final %>% filter(supergroup_boot > 90)

  otu_mat<- otu_table_euk_final %>%  select(otu=otu_id,matches("EC|PR|RM|SBW|STJ"),-species, -species_boot )
  tax_mat<- otu_table_euk_final %>%  select(otu=otu_id,kingdom:species)
  samples_df <- sample_table %>%  rename (sample=sample_id)
  
  row.names(otu_mat) <- otu_mat$otu
  otu_mat <- otu_mat %>% select (-otu) 
  row.names(tax_mat) <- tax_mat$otu
  tax_mat <- tax_mat %>% select (-otu) 
  row.names(samples_df) <- samples_df$sample
  samples_df <- samples_df %>% select (-sample)
  otu_mat <- as.matrix(otu_mat)
  tax_mat <- as.matrix(tax_mat)
  
  OTU = otu_table(otu_mat, taxa_are_rows = TRUE)
  TAX = tax_table(tax_mat)
  samples = sample_data(samples_df)
  
  singa_euk <- phyloseq(OTU, TAX, samples)
  singa_euk <- subset_samples(singa_euk, sequence_quality == "good")

  cat("\nPhyloseq Eukaryotes \n========== \n")  
  singa_euk
```
## Break up into photosynthetic and non-photosynthetic 
* **Opisthokonta (Metazoa, Fungi) are removed**

```{r}
  singa_euk <- subset_taxa(singa_euk,!(supergroup %in% c("Opisthokonta")))

  singa_photo <- subset_taxa(singa_euk, (division %in% c("Chlorophyta", "Cryptophyta", "Rhodophyta",
                                                            "Haptophyta", "Ochrophyta")) |
                                        ((division ==  "Dinoflagellata") & (class != "Syndiniales")) |
                                        (class == "Filosa-Chlorarachnea"))
  cat("\nPhyloseq Photosynthetic Eukaryotes \n========== \n") 
  singa_photo
  
  
  singa_hetero <- subset_taxa(singa_euk, !(division %in% c("Chlorophyta", "Cryptophyta", "Rhodophyta",
                                                            "Haptophyta", "Ochrophyta")) &
                                        !((division ==  "Dinoflagellata") & !(class == "Syndiniales")) &
                                        !(class == "Filosa-Chlorarachnea"))
  cat("\nPhyloseq Heterotrophic Eukaryotes \n========== \n") 
  singa_hetero
```

## Create phyloseq files for proks

```{r}
  otu_table_prok <- otu_table %>% filter(taxo1 %in% c("Bacteria", "Archaea"))
  
  otu_mat<- otu_table_prok %>%  select(otu=otu_id,matches("EC|PR|RM|SBW|STJ") )
  
  tax_mat<- otu_table_prok %>%  select(otu=otu_id,taxo1:taxo7) %>% 
    rename(kingdom=taxo1, supergroup=taxo2, division=taxo3, class=taxo4, order=taxo5, family=taxo6, genus=taxo7) %>% 
    mutate(species=NA)
  
  samples_df <- sample_table %>%  rename (sample=sample_id)
  
  row.names(otu_mat) <- otu_mat$otu
  otu_mat <- otu_mat %>% select (-otu) 
  row.names(tax_mat) <- tax_mat$otu
  tax_mat <- tax_mat %>% select (-otu) 
  row.names(samples_df) <- samples_df$sample
  samples_df <- samples_df %>% select (-sample)
  otu_mat <- as.matrix(otu_mat)
  tax_mat <- as.matrix(tax_mat)
  
  OTU = otu_table(otu_mat, taxa_are_rows = TRUE)
  TAX = tax_table(tax_mat)
  samples = sample_data(samples_df)
  
  singa_prok <- phyloseq(OTU, TAX, samples)
  singa_prok <- subset_samples(singa_prok, sequence_quality == "good")
  
  cat("\nPhyloseq Prokaryotes \n========== \n") 
  singa_hetero
```

## Normalize number of reads in each sample using median sequencing depth.
* ! If there no cells do not transform, just set column to 0 
`function(x, t=total_hetero) (if(sum(x) > 0){ t * (x / sum(x))} else {x})`


```{r}

# First define a function to normalize
   
ps_normalize_median <- function (ps, title) {
  ps_median = median(sample_sums(ps))
  cat(sprintf("\nThe median number of reads used for normalization of %s is  %.0f", title, ps_median))
  normalize_median = function(x, t=ps_median) (if(sum(x) > 0){ t * (x / sum(x))} else {x})
  ps = transform_sample_counts(ps, normalize_median)
  cat(str_c("\nPhyloseq ",title, "\n========== \n") )
  print(ps)
}

# Apply to all the phyloseq files
  singa_euk = ps_normalize_median(singa_euk, "eukaryotes (auto+hetero)")
  singa_photo = ps_normalize_median(singa_photo, "eukaryotes autotrophs")
  singa_hetero = ps_normalize_median(singa_hetero, "eukaryotes heterotrophs")
  singa_prok = ps_normalize_median(singa_prok, "prokaryotes")
  
```

## Remove taxa which are in low abundance
* Remove taxa that are < 0.10 (euks) and <0.05 (proks) in any given sample
* Normalize again...

```{r}

ps_abundant <- function(ps,contrib_min=0.1, title){
     total_per_sample <- max(sample_sums(ps))
     ps <- filter_taxa(ps, function(x) sum(x > total_per_sample*contrib_min) > 0, TRUE)
     ps <- ps_normalize_median(ps, title)
}

  cat("Remove taxa in low abundance \n\n")
  singa_euk_abundant = ps_abundant(singa_euk,contrib_min=0.1, "eukaryotes (auto+hetero)")
  singa_photo_abundant = ps_abundant(singa_photo,contrib_min=0.1, "eukaryotes autotrophs")
  singa_hetero_abundant = ps_abundant(singa_hetero,contrib_min=0.1, "eukaryotes heterotrophs")
  singa_prok_abundant = ps_abundant(singa_prok,contrib_min=0.05, "prokaryotes")

```


## Create a list for the auto and hetero phyloseq files

```{r}
  ps_list <- list(ps=c(singa_prok,  singa_euk, singa_photo ), 
                  title = c( "Prokaryotes - all OTUs", 
                             "Eukaryotes - Auto and Hetero - all OTUs", 
                             "Eukaryotes - Autotrophs - all OTUs"))
  ps_list_abundant <- list(ps=c( singa_prok_abundant, singa_euk_abundant, singa_photo_abundant), 
                           title = c( "Prokaryotes - abundant OTUs (> 5%)", 
                                      "Eukaryotes - Auto + Hetero - abundant OTUs (> 10%)",
                                      "Eukaryotes - Autotrophs - abundant OTUs (> 10%)"))
```


## Bar plot of divisions per station


Note: some stations are completely missing heterotrops (Only Opistokonta)

```{r fig.height=15, fig.width=10}

for (i in 1:3) {
  p <- plot_bar(ps_list$ps[[i]], x="sample_label", fill = "division") + 
       geom_bar(aes(color=division, fill=division), stat="identity", position="stack") + 
       ggtitle(str_c("Division level - ",ps_list$title[[i]]))  +
       theme(axis.text.y = element_text(size=10)) +
       theme(axis.text.x = element_text(angle=0, hjust = 0.5)) +
       coord_flip()
  print(p)
}
```

## Bar plot of class per station

Only consider the abundant taxa

```{r fig.height=15, fig.width=10}

for (i in 1:3) {
  p <- plot_bar(ps_list_abundant$ps[[i]], x="sample_label", fill = "class") + 
       geom_bar(aes(color=class, fill=class), stat="identity", position="stack") + 
       ggtitle(str_c("Class level - ",ps_list_abundant$title[[i]])) +
       theme(axis.text.y = element_text(size=10))+
       theme(axis.text.x = element_text(angle=0, hjust = 0.5)) +
       coord_flip() 
  print(p)
}
```

## Compare by Straight, Site, Moonsoon (abundant OTUs only)

```{r fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
for (factor in c("strait", "location", "monsoon")) {  
 for (i in 1:3) {
  ps_aggregate <- merge_samples(ps_list_abundant$ps[[i]], factor)
  p <- plot_bar(ps_aggregate, fill = "division") + 
      geom_col(aes(color=division, fill=division)) +
      ggtitle(str_c(ps_list_abundant$title[[i]], " - ", factor)) +
      theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1))
  print(p)
 }
}
```

## Main taxa

### Main genera for different division of Eukaryotes (Autotrophs)

```{r fig.height=6, fig.width=8}
for (one_division in c("Chlorophyta", "Dinoflagellata", "Ochrophyta")){
  ps_subset <- subset_taxa(singa_photo_abundant, division %in% one_division)
  p <- plot_bar(ps_subset, x="genus",  facet_grid = strait~monsoon) +
    geom_col() +
    theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
    ggtitle(str_c(one_division," - Abundant OTUs")) +
    coord_flip()
  print(p)  
}
```

### Main species of Mamiellophyceae and Diatoms (Eukaryotes - Autrotrophs)

```{r fig.height=6, fig.width=8}
for (one_class in c("Mamiellophyceae", "Bacillariophyta")){
  ps_subset <- subset_taxa(singa_photo_abundant, class %in% one_class)
  p <- plot_bar(ps_subset, x="species", facet_grid = strait~monsoon) +
    geom_col() +
    theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
    ggtitle(str_c(one_class," - Abundant OTUs"))+
    coord_flip()
  print(p)  
}
```

## Heatmaps (abundant OTUs)

* Data are agglomated at the genus level. Use function `tax_glom`

```{r fig.height=10, fig.width=10}

for (i in c(1)) {
  ps_heat <- tax_glom(ps_list_abundant$ps[[i]],taxrank = "family")
  p <- plot_heatmap(ps_heat, method = "NMDS", distance = "bray", 
               taxa.label = "family",  taxa.order="division",
               sample.label = "sample_label", sample.order="sample_label",
               low="beige", high="red", na.value="beige", 
               title=ps_list_abundant$title[[i]])
  print(p)
}
  

for (i in 2:3) {
  
  ps_heat <- tax_glom(ps_list_abundant$ps[[i]],taxrank = "genus")
  p <- plot_heatmap(ps_heat, method = "NMDS", distance = "bray", 
               taxa.label = "genus",  taxa.order="division",
               sample.label = "sample_label", sample.order="sample_label",
               low="beige", high="red", na.value="beige", 
               title=ps_list_abundant$title[[i]])
  print(p)
}


```

## NMDS



Sample removed because they were pulling the NMDS
* PR2X16XS21 it has a single eukaryote (diatom bloom )
* RM13XS36 cause problem for bacteria
* PR11XS25 cause problem for hetero euks
* SBW11XS26 cause problem for hetero euks
* SBW13XS37 cause problem for hetero euks
* RM13XS36 cause problem for hetero euks


Define function

```{r}
singa_msds <- function(ps_list) {
    for (i in 1:3) {
      ps_nmds <- ps_list$ps[[i]]
      
      # Remove samples with no reads
      ps_nmds <- prune_samples(sample_sums(ps_nmds)>0, ps_nmds)
      
      # Remove samples that caused problems (1= prok, 2=euk, 3=euk auto)
      if (i==1) {  ps_nmds <- prune_samples(!(sample_names(ps_nmds) %in% c("RM13XS36")), ps_nmds) } # Prokaryotes
      if (i %in% c(2,3)) {  ps_nmds <- prune_samples(!(sample_names(ps_nmds) %in% c("PR2X16SXS21")), ps_nmds)} # Eukaryotes
      if (i==4) {  ps_nmds <- prune_samples(!(sample_names(ps_nmds) %in% c("PR11XS25", "SBW11XS26", "SBW13XS37")), ps_nmds) } # Heterotrophs not used
       
      
      
      singa.ord <- ordinate(ps_nmds, "NMDS", "bray")
      print(singa.ord)
    
      p <- plot_ordination(ps_nmds, singa.ord, type="samples", color="strait", 
                      shape="monsoon", title=ps_list$title[[i]]) +
           geom_point(size=5)  +
           scale_color_manual(values=c("red", "green","blue")) +
           scale_shape_manual(values = c(17,18,15,16))
           theme_bw()
    
      print(p)
      
       p <- plot_ordination(ps_nmds, singa.ord, type="taxa", color="supergroup", 
                  shape="supergroup", title=ps_list$title[[i]]) +
       geom_point(size=3) +
       theme_bw()
  print(p)
    }
  }
```

### All OTUs 

```{r, fig.height=8, fig.width=8}

  singa_msds(ps_list)
  
```

### Abundant OTUs
```{r, fig.height=8, fig.width=8}
  singa_msds(ps_list_abundant)

```

## Network analysis

```{r fig.height=6, fig.width=10}

for (i in 1:3) {
  
  ps_nmds <- ps_list_abundant$ps[[i]]
  
  # Remove samples with no reads
  ps_nmds <- prune_samples(sample_sums(ps_nmds)>0, ps_nmds)
  
  # Remove samples that caused problems (1= prok, 2=euk, 3=euk auto)
  if (i==1) {  ps_nmds <- prune_samples(!(sample_names(ps_nmds) %in% c("RM13XS36")), ps_nmds) } # Prokaryotes
  if (i %in% c(2,3)) {  ps_nmds <- prune_samples(!(sample_names(ps_nmds) %in% c("PR2X16SXS21")), ps_nmds)} # Eukaryotes
  if (i==4) {  ps_nmds <- prune_samples(!(sample_names(ps_nmds) %in% c("PR11XS25", "SBW11XS26", "SBW13XS37")), ps_nmds) } # Heterotrophs not used

  if (i > 1) {
  p <- plot_net(ps_nmds, distance = "(A+B-2*J)/(A+B)", type = "taxa", 
           maxdist = 0.4, color="class", point_label="genus") +
       ggtitle(ps_list_abundant$title[[i]])
  }

  else {
  p <- plot_net(ps_nmds, distance = "(A+B-2*J)/(A+B)", type = "taxa", 
           maxdist = 0.4, color="class", point_label="family") +
       ggtitle(ps_list_abundant$title[[i]])
  }

  print(p)
}

```



