---
title: "Metabarcoding Singapore <br> Phyloseq"
author: "Daniel Vaulot, Adriana Lopes dos Santos"
date: '`r format(Sys.time(), "%d %m %Y")`'
header-includes:
   - \usepackage{color, fancyvrb}
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    number_sections : yes    
  pdf_document: 
    toc: yes
    toc_depth: 3
    number_sections : yes
---


# Aim

* Create Phyloseq file from samples, metadata and OTU table (read from Excel file)

* https://bioconductor.org/help/course-materials/2017/BioC2017/Day1/Workshops/Microbiome/MicrobiomeWorkflowII.html#abstract

# Initialize

This file defines all the necessary libraries and variables

```{r init, eval=TRUE, message=FALSE, warning=FALSE}
  source('Metabarcoding Singapore_init.R', echo=FALSE)
```

# Read the data
## File names

```{r}
# Metadata and stations
  metadata_xlsx <- "Singapore_metadata.xlsx"
  metadata_csv <- "monsoonpaper_env_data.csv"
  
# OTUs
  otu_excel_file <- "Singapore ASV_table.xlsx"
```


## Read the files
* The dada2 treatment has already removed the forward and reverse primers, so no need to remove them
* Work with the unrarefied data

### Metadata and sample files

```{r}
# Read the sample and metadata tables
  sample_table <- read_excel(metadata_xlsx, sheet="samples", range="A1:D89") 

  metadata_table  <- read_csv(metadata_csv, na=c("ND", "")) %>% 
    dplyr::rename(sample_code=Sample, day_number=Day_number, date=Date, 
                  location=Location, monsoon=`Monsoon period`) %>% 
    select(-Strait) %>% 
    mutate(date = lubridate::parse_date_time(date,"dmy"),
           monsoon = forcats::fct_relevel(monsoon, "NE", "IM-1", "SW", "IM-2"))
  
  station_table  <- read_excel(metadata_xlsx, sheet="stations", na=c("ND", ""))

  sample_table <- sample_table %>% 
    left_join(metadata_table) %>%
    left_join(station_table) %>% 
    mutate(sample_label = str_c(location_label,
                                date,
                                monsoon, 
                                sep="_"))

```
### OTU table
NOTE: The OTU table is read from Excel file

The following editing has been done
* Nanochloris changed to Picochlorum based on BLAST table
* Ostreococcus sp. changed to O. lucimarinus based on the BLAST


```{r}
# Read the OTU table

  otu_table_final <- read_excel(otu_excel_file, sheet="ASV")

```


# Map
## Leaflet map

* Visualize different layers: https://leaflet-extras.github.io/leaflet-providers/preview/

```{r eval=TRUE, fig.height=10, fig.width=10}

  lng_center = mean(station_table$longitude)
  lat_center = mean(station_table$latitude)

  map <- leaflet(width = 1000, height = 1000) %>%
   addTiles() %>%  # Default
#    addTiles(urlTemplate = 'https://server.arcgisonline.com/ArcGIS/rest/services/World_Imagery/MapServer/tile/{z}/{y}/{x}' ) %>%  # Satellite
#    addTiles(urlTemplate = 'https://server.arcgisonline.com/ArcGIS/rest/services/Canvas/World_Light_Gray_Base/MapServer/tile/{z}/{y}/{x}' ) %>%  # Grey background
    setView(lng=lng_center, lat=lat_center, zoom=11) %>% 
    addCircleMarkers(data = station_table, lat = ~ latitude, lng = ~ longitude, 
                     radius = 5,
               label = ~ location, 
               labelOptions = labelOptions(textsize = "10px", noHide = T),
               clusterOptions = markerClusterOptions()) 

  map
```


# Environmental data

## Per station
### Temp
```{r temp_per_station, fig.height=10, fig.width=8}
 p <- ggplot(filter(sample_table, location_label != "BL")) + 
  geom_line(aes(x=date,y=Temperature), na.rm=TRUE) +
  geom_point(aes(x=date,y=Temperature, color=monsoon), size=5) +
  facet_grid(rows = vars(location)) +
  ylim(26,34) +
  xlim(as.POSIXct(as.Date(c("2017-01-01","2019-01-01" )))) + 
  geom_vline(xintercept = as.POSIXct(as.Date(c("2017-01-01","2018-01-01", "2019-01-01")))) +
  scale_color_viridis(discrete=TRUE) 
 p
 

```
### Salinity
```{r salinity_per_station, fig.height=10, fig.width=8}
 p <- ggplot(filter(sample_table, location_label != "BL")) + 
  geom_line(aes(x=date,y=Salinity), na.rm=TRUE) +
  geom_point(aes(x=date,y=Salinity, color=monsoon), size=5) +
  facet_grid(rows = vars(location)) +
  ylim(15,35) +
  xlim(as.POSIXct(as.Date(c("2017-01-01","2019-01-01" )))) + 
  geom_vline(xintercept = as.POSIXct(as.Date(c("2017-01-01","2018-01-01", "2019-01-01")))) +
  scale_color_viridis(discrete=TRUE) 
 p
 

```

### Chlorophyll

```{r chl_per_station, fig.height=10, fig.width=8}
 p <- ggplot(filter(sample_table, location_label != "BL")) + 
  geom_line(aes(x=date,y=Chl), na.rm=TRUE) +
  geom_point(aes(x=date,y=Chl, color=monsoon), size=5) +
  facet_grid(rows = vars(location)) +
  xlim(as.POSIXct(as.Date(c("2017-01-01","2019-01-01" )))) + 
  geom_vline(xintercept = as.POSIXct(as.Date(c("2017-01-01","2018-01-01", "2019-01-01")))) +
  scale_color_viridis(discrete=TRUE) 
 p
 

```

### Rain over 7 days

```{r rain_per_station, fig.height=10, fig.width=8}
 p <- ggplot(filter(sample_table, location_label != "BL")) + 
  geom_col(aes(x=date,y=Rain7, fill=monsoon)) + 
  facet_grid(rows = vars(location)) +
  xlim(as.POSIXct(as.Date(c("2017-01-01","2019-01-01" )))) + 
  geom_vline(xintercept = as.POSIXct(as.Date(c("2017-01-01","2018-01-01", "2019-01-01")))) +
  scale_fill_viridis(discrete=TRUE) 
 p
 

```

## Average
### Rain over 7 days
```{r rain_mean_per_station}
 rain <- sample_table %>% 
   group_by(date, monsoon) %>%
   summarise(Rain7_mean = mean(Rain7, na.rm=TRUE))
 
 p <- ggplot(rain) + 
  geom_point(aes(x=date,y=Rain7_mean, color=monsoon), size=5) +
  xlim(as.POSIXct(as.Date(c("2017-01-01","2019-01-01" )))) + 
  geom_vline(xintercept = as.POSIXct(as.Date(c("2017-01-01","2018-01-01", "2019-01-01")))) +
  scale_color_viridis(discrete=TRUE) 
 p
 
```


# Phyloseq analysis

## Create phyloseq files for all taxa after filtering the data

Filter the euk data to remove the low bootstraps values (threshold : bootstrap > 90% at the supergroup level) and create a phyloseq file

Note the bootstrap threshold 90% for ASVs 



### Filter the ASV table to remove low bootstraps
```{r}
  otu_table_euk_final <- otu_table_final %>% filter(supergroup_boot > 90)
  otu_table_prok_final <- otu_table_final %>% filter(kingdom %in% c("Bacteria", "Archaea")) 
  
  otu_table_ps <- bind_rows(otu_table_euk_final, otu_table_prok_final) %>% 
    arrange(otu_id)
```

### Create the phyloseq file
* Remove sequences with bad quality
* Remove Metazoa
* Remove Chloroplast


```{r warning=FALSE}
  otu_mat<- otu_table_ps %>%  select(otu=otu_id,matches("EC|PR|RM|SBW|STJ"),-species, -species_boot )
  tax_mat<- otu_table_ps %>%  select(otu=otu_id,kingdom:species)
  samples_df <- sample_table %>%  rename (sample=sample_id)
  
  row.names(otu_mat) <- otu_mat$otu
  otu_mat <- otu_mat %>% select (-otu) 
  row.names(tax_mat) <- tax_mat$otu
  tax_mat <- tax_mat %>% select (-otu) 
  row.names(samples_df) <- samples_df$sample
  samples_df <- samples_df %>% select (-sample)
  otu_mat <- as.matrix(otu_mat)
  tax_mat <- as.matrix(tax_mat)
  
  OTU = otu_table(otu_mat, taxa_are_rows = TRUE)
  TAX = tax_table(tax_mat)
  samples = sample_data(samples_df)
  
  ps_all_raw <- phyloseq(OTU, TAX, samples)
  
# Filtering
  ps_all_raw <- subset_samples(ps_all_raw, sequence_quality == "good")
  ps_all_raw <- subset_taxa(ps_all_raw,!(division %in% c("Metazoa")))
  ps_all_raw <- subset_taxa(ps_all_raw,!(kingdom %in% c("Eukaryota_plastid")))  

```
## Break up into photosynthetic and non-photosynthetic 


```{r}
  cat("\nPhyloseq All \n========== \n")  
  ps_all <- ps_all_raw
  ps_all

  ps_euk <- subset_taxa(ps_all,(kingdom %in% c("Eukaryota")))
  cat("\nPhyloseq Eukaryotes \n========== \n")  
  ps_euk

  ps_photo <- subset_taxa(ps_euk, (division %in% c("Chlorophyta", "Cryptophyta", "Rhodophyta",
                                                            "Haptophyta", "Ochrophyta")) |
                                        ((division ==  "Dinoflagellata") & (class != "Syndiniales")) |
                                        (class == "Filosa-Chlorarachnea"))
  cat("\nPhyloseq Photosynthetic Eukaryotes \n========== \n") 
  ps_photo
  
  
  ps_hetero <- subset_taxa(ps_euk, !(division %in% c("Chlorophyta", "Cryptophyta", "Rhodophyta",
                                                            "Haptophyta", "Ochrophyta")) &
                                        !((division ==  "Dinoflagellata") & !(class == "Syndiniales")) &
                                        !(class == "Filosa-Chlorarachnea"))
  cat("\nPhyloseq Heterotrophic Eukaryotes \n========== \n") 
  ps_hetero
  
  ps_prok <- subset_taxa(ps_all,(kingdom %in% c("Bacteria", "Archaea")))
  cat("\nPhyloseq Prokaryotes \n========== \n")  
  ps_prok
  
```


## Normalize number of reads in each sample using median sequencing depth.
* ! If there no cells do not transform, just set column to 0 
`function(x, t=total_hetero) (if(sum(x) > 0){ t * (x / sum(x))} else {x})`


```{r}

# First define a function to normalize
   
ps_normalize_median <- function (ps, title) {
  ps_median = median(sample_sums(ps))
  cat(sprintf("\nThe median number of reads used for normalization of %s is  %.0f", title, ps_median))
  normalize_median = function(x, t=ps_median) (if(sum(x) > 0){ t * (x / sum(x))} else {x})
  ps = transform_sample_counts(ps, normalize_median)
  cat(str_c("\nPhyloseq ",title, "\n========== \n") )
  print(ps)
}

# Apply to all the phyloseq files
  ps_all = ps_normalize_median(ps_all, "all")
  ps_euk = ps_normalize_median(ps_euk, "eukaryotes (auto+hetero)")
  ps_photo = ps_normalize_median(ps_photo, "eukaryotes autotrophs")
  ps_hetero = ps_normalize_median(ps_hetero, "eukaryotes heterotrophs")
  ps_prok = ps_normalize_median(ps_prok, "prokaryotes")
  
```

## Phyloseq files for abundant taxa
* Remove taxa that are < 0.10 (euks) and <0.05 (proks) in any given sample
* Normalize again...

```{r}

ps_abundant <- function(ps,contrib_min=0.1, title){
     total_per_sample <- max(sample_sums(ps))
     ps <- filter_taxa(ps, function(x) sum(x > total_per_sample*contrib_min) > 0, TRUE)
     ps <- ps_normalize_median(ps, title)
}

  cat("Remove taxa in low abundance \n\n")
  ps_all_abundant = ps_abundant(ps_all,contrib_min=0.05, "All")  
  ps_euk_abundant = ps_abundant(ps_euk,contrib_min=0.1, "eukaryotes (auto+hetero)")
  ps_photo_abundant = ps_abundant(ps_photo,contrib_min=0.1, "eukaryotes autotrophs")
  ps_hetero_abundant = ps_abundant(ps_hetero,contrib_min=0.1, "eukaryotes heterotrophs")
  ps_prok_abundant = ps_abundant(ps_prok,contrib_min=0.05, "prokaryotes")

```


## Create a list for the auto and hetero phyloseq files

```{r}
  ps_list <- list(ps=c(ps_prok,  ps_euk, ps_photo ), 
                  title = c( "Prokaryotes - all OTUs", 
                             "Eukaryotes - Auto and Hetero - all OTUs", 
                             "Eukaryotes - Autotrophs - all OTUs"))
  ps_list_abundant <- list(ps=c( ps_prok_abundant, ps_euk_abundant, ps_photo_abundant), 
                           title = c( "Prokaryotes - abundant OTUs (> 5%)", 
                                      "Eukaryotes - Auto + Hetero - abundant OTUs (> 10%)",
                                      "Eukaryotes - Autotrophs - abundant OTUs (> 10%)"))
```

## Create tabular files for other plots

```{r}
  ps_to_long <- function(ps) {
    otu_df <- data.frame(otu_table(ps)) %>% 
      rownames_to_column(var = "otu_id")
    taxo_df <- data.frame(tax_table(ps))%>% 
      rownames_to_column(var = "otu_id")
    otu_df <- left_join(taxo_df, otu_df)
    otu_df <- gather(otu_df, "sample","n_seq" ,contains("X")) # All samples contain X
    metadata_df <- data.frame(sample_data(ps)) %>% 
      rownames_to_column(var = "sample")
    otu_df <- left_join(otu_df, metadata_df)
  }

  long_all <- ps_to_long(ps_all)
  long_euk <- ps_to_long(ps_euk) 
  long_photo <- ps_to_long(ps_photo)
  long_euk_abundant <- ps_to_long(ps_euk_abundant)
  long_photo_abundant <- ps_to_long(ps_photo_abundant)
  long_prok <- ps_to_long(ps_prok)
```


## Treemaps at division and class levels


### Define function to draw treemaps

```{r}
 treemap_gg_dv <- function(df, group1, group2, title) {
     group1 <- enquo(group1)
     group2 <- enquo(group2)
     df <- df %>% group_by(!!group1, !!group2) %>%  summarise(n_seq=sum(n_seq))
     g_treemap <- ggplot(df, aes(area = n_seq, fill = !!group2, 
                                              label = !!group2, subgroup = !!group1)) +
        ggtitle(title) +
        treemapify::geom_treemap() +
        treemapify::geom_treemap_subgroup_border() +
        treemapify::geom_treemap_text(colour = "black", place = "topleft", reflow = T, 
                                      padding.x =  grid::unit(3, "mm"), 
                                      padding.y = grid::unit(3, "mm") ) +
        treemapify::geom_treemap_subgroup_text(place = "centre", grow = T, alpha = 0.5, colour =
                             "white", fontface = "italic", min.size = 0) +
        scale_fill_viridis(discrete=TRUE) +
        theme(legend.position="none", plot.title = element_text(size = 16, face = "bold"))
     print(g_treemap)
     return(g_treemap)
}
```

### Do the individual treemaps

```{r treemap, fig.height=8, fig.width=8}




  array_treemap_all <- list()
  array_treemap_by_kingdom <- list()
  
  for (one_strait in c("Singapore", "Johor")) {
   
  label <- str_c(one_strait,"-all") 

  array_treemap_all[[label]] <- treemap_gg_dv(filter(long_all, strait == one_strait), 
                                 kingdom, supergroup, str_c("", one_strait )) 

  label <- str_c(one_strait,"-arch") 
  array_treemap_by_kingdom[[label]] <- treemap_gg_dv(filter(long_prok, strait == one_strait & kingdom=="Archaea"),
                                          division, class, str_c("A - Archaea - ", one_strait ))
  
  label <- str_c(one_strait,"-bact") 
  array_treemap_by_kingdom[[label]] <- treemap_gg_dv(filter(long_prok, strait == one_strait & kingdom=="Bacteria"),
                                          division, class, str_c("B - Bacteria - ", one_strait ))
  
     
  label <- str_c(one_strait,"-euks") 
  array_treemap_by_kingdom[[label]] <- treemap_gg_dv(filter(long_euk, strait == one_strait), 
                                          division, class, str_c("C - All Eulkaryotes - ", one_strait )) 
  
   
  # label <- str_c(one_strait,"-photeuks") 
  # array_treemap[[label]] <- treemap_gg_dv(filter(long_photo, strait == one_strait), 
  #                                         division, class, str_c("D - Photosynthetic Eukaryotes - ", one_strait ))
  # 
   

  # treemap_dv(filter(long_euk, strait == one_strait), 
  #            c("division", "class"),"n_seq", str_c("All euks - ", one_strait ))
  # treemap_dv(filter(long_photo, strait == one_strait),
  #            c("division", "class"),"n_seq",str_c("Photo euks", one_strait ))
  # treemap_dv(filter(long_prok, strait == one_strait & kingdom=="Bacteria"),
  #            c("division", "class"),"n_seq",str_c("Bacteria - ", one_strait ))
  # treemap_dv(filter(long_prok, strait == one_strait & kingdom=="Archaea"),
  #            c("division", "class"),"n_seq",str_c("Archaea - ", one_strait ))
  }


```

### FIG - Arrange the different treemaps in a grid and save

```{r eval=TRUE, fig.height=10, fig.width=5}
# Main Fig, all groups
fig_grid_treemap_all <- gridExtra::grid.arrange(grobs=array_treemap_all, 
                                                ncol=2, nrow=1, clip=FALSE, 
                                                padding = unit(0, "line"),
                                                as.table = FALSE) 

fig_grid_treemap_all

ggsave("../fig/fig_treemap_all.png", fig_grid_treemap_all, height = 5, width = 10, dpi = 300)


# Supp Fig, by kingdom

fig_grid_treemap_by_kingdom <- gridExtra::grid.arrange(grobs=array_treemap_by_kingdom, 
                                                      ncol=2, nrow=3, clip=FALSE, 
                                                      padding = unit(0, "line"),
                                                      as.table = FALSE) 

fig_grid_treemap_by_kingdom

ggsave("../fig/fig_treemap_separate.png", fig_grid_treemap_by_kingdom, height = 15, width = 10, dpi = 300)
   
```

## Most abundant taxa

### Most abundant eukaryote species

```{r abundant_species, fig.height=6, fig.width=8}
  df <- filter(long_euk, strait %in% c("Johor", "Singapore")) %>%
    mutate(species_label = str_c(class, species, sep="-")) %>% 
    group_by(division, class, species, species_label, strait) %>% 
    summarize(n_seq = sum(n_seq)) %>% 
    arrange(desc(n_seq)) %>%  
    filter(n_seq > 0) %>% 
    ungroup()

  df_strait <- df %>%
    select(species, strait) %>% 
    group_by(species) %>% 
    summarise(n_strait = n()) %>% 
    ungroup()

  df <- df %>% 
    left_join(df_strait, by = c("species"="species")) %>% 
    mutate(species = case_when(
                            n_strait == 1 ~ str_c(species, " *"),
                            TRUE  ~  species
                          ) )

  array_euk <- list()


  for (one_strait in c("Singapore", "Johor")) {

  df_one <- filter(df, strait == one_strait) 

  array_euk[[one_strait]] <- ggplot(top_n(df_one, 30, n_seq)) + 
    geom_col(aes(x=reorder(species, n_seq), y=n_seq, fill=division)) + 
    coord_flip() +
    xlab("") + ylab("Number of reads") +
    scale_fill_manual(values = division_euk_colors) +
    ggtitle(one_strait) +
    theme_bw() +
    theme(plot.title = element_text(size=22, hjust = 0.5)) +
    # theme(axis.text=element_text(size=14), legend.text = element_text(size=16)) + 
    theme(legend.position = "top", legend.box = "vertical") +
    guides(fill = guide_legend(title.position="top", 
                               ncol = 4, nrow = 2, byrow = FALSE))
  
  print(array_euk[[one_strait]])
  
  }


```

### Most abundant prokaryotes taxo6 (~ genus)

```{r abundant_genus_prok, fig.height=6, fig.width=10}
  df <- filter(long_prok, strait %in% c("Johor", "Singapore")) %>%
    mutate(genus_label = str_c(division, family, sep="-")) %>% 
    group_by(supergroup, division, class, family, genus_label, strait) %>% 
    summarize(n_seq = sum(n_seq)) %>% 
    arrange(desc(n_seq)) %>%  
    filter(n_seq > 0) %>% 
    ungroup()

  df_strait <- df %>%
    select(genus_label, strait) %>% 
    group_by(genus_label) %>% 
    summarise(n_strait = n()) %>% 
    ungroup()

  df <- df %>% 
    left_join(df_strait) %>% 
    mutate(genus_label = case_when(
                            n_strait == 1 ~ str_c(genus_label, " *"),
                            TRUE  ~  genus_label
                          ) )


  array_prok <- list()

  for (one_strait in c("Singapore", "Johor")) {

  df_one <- filter(df, strait == one_strait) 

  array_prok[[one_strait]] <- ggplot(top_n(df_one, 30, n_seq)) + 
    geom_col(aes(x=reorder(genus_label, n_seq), y=n_seq, fill=supergroup)) + 
    coord_flip() +
    xlab("") + ylab("Number of reads") +
    scale_fill_manual(values = supergroup_colors) +
    ggtitle(one_strait) +
    theme_bw() +
    theme(plot.title = element_text(size=22, hjust = 0.5)) +
    # theme(axis.text=element_text(size=14), legend.text = element_text(size=16)) + 
    theme(legend.position = "top", legend.box = "vertical") +
    guides(fill = guide_legend(title.position="top", 
                               ncol = 4, nrow = 2, byrow = FALSE))
  
  print(array_prok[[one_strait]])
  
  }


```

### Most abundant taxa - Save as Grid
```{r}
   grid_fig <-  cowplot::plot_grid(array_prok[["Singapore"]],array_prok[["Johor"]], 
                                   array_euk[["Singapore"]],array_euk[["Johor"]], align="hv",
                                   # labels=c("A","B", "C", "D"),
                                   ncol=2, nrow=2, label_size = 20)


   # ggsave("../fig/fig_top_all.png", grid_fig, 
   #        height = 20, width = 20, dpi = 300)

```

## Most abundant OTUs

### Compute and plot separately Euk and Prok

```{r abundant_OTUS, fig.height=12, fig.width=12}
  df <- filter(long_all, strait %in% c("Johor", "Singapore")) %>%
    mutate(otu_label = case_when (kingdom == "Eukaryota" ~ str_c(otu_id, species, sep="-"),
                                  TRUE ~  str_c(otu_id, family, sep="-") )) %>% 
    group_by(kingdom, supergroup, division, class, genus, species, otu_id, otu_label, strait) %>% 
    summarize(n_seq = sum(n_seq)) %>% 
    arrange(desc(n_seq)) %>%  
    filter(n_seq > 0) %>% 
    ungroup()

  df_strait <- df %>%
    select(otu_label, strait) %>% 
    group_by(otu_label) %>% 
    summarise(n_strait = n()) %>% 
    ungroup()

  df <- df %>% 
    left_join(df_strait) %>% 
    mutate(otu_label = case_when(
                            n_strait == 1 ~ str_c(otu_label, " *"),
                            TRUE  ~  otu_label
                          ) )
  
  otu_strait <- df
  
  array_euk <- list()
  array_prok <- list()
  
# Do plots

  for (one_strait in c("Singapore", "Johor")) {

# For euks

  df_one <- filter(df, strait == one_strait & kingdom == "Eukaryota") 

  array_euk[[one_strait]] <- ggplot(top_n(df_one, 30, n_seq)) + 
    geom_col(aes(x=reorder(otu_label, n_seq), y=n_seq, fill=division)) + 
    coord_flip() +
    theme_bw() +
    theme(axis.text.x = element_text(size = 16, angle = 0, hjust = 1, vjust = 1)) +
    theme(axis.text.y = element_text(size = 16, angle = 0, hjust = 0, vjust = 0)) +
    theme(legend.title = element_text(size = 24)) +
    theme(legend.text = element_text(size = 16)) +
    xlab("") + ylab("Number of reads") +
    scale_fill_manual(values = division_euk_colors) +
    ggtitle(one_strait)  +
    theme(plot.title = element_text(size=22, hjust = 0.5)) +
    # theme(axis.text=element_text(size=14), legend.text = element_text(size=16)) + 
    theme(legend.position = "top", legend.box = "vertical") +
    guides(fill = guide_legend(title.position="top", 
                               ncol = 2, nrow = 4, byrow = FALSE))
  
  print(array_euk[[one_strait]])
  
# For proks

  
  df_one <- filter(df, strait == one_strait & kingdom != "Eukaryota") 

  array_prok[[one_strait]] <- ggplot(top_n(df_one, 30, n_seq)) + 
    geom_col(aes(x=reorder(otu_label, n_seq), y=n_seq, fill=supergroup)) + 
    coord_flip() +
    theme_bw() +
    theme(axis.text.x = element_text(size = 16, angle = 0, hjust = 1, vjust = 1)) +
    theme(axis.text.y = element_text(size = 16, angle = 0, hjust = 0, vjust = 0)) +
    theme(legend.title = element_text(size = 24)) +
    theme(legend.text = element_text(size = 16)) +
    xlab("") + ylab("Number of reads") +
    scale_fill_manual(values = supergroup_colors) +
    ggtitle(one_strait)  +
    theme(plot.title = element_text(size=22, hjust = 0.5)) +
    # theme(axis.text=element_text(size=14), legend.text = element_text(size=16)) + 
    theme(legend.position = "top", legend.box = "vertical") +
    guides(fill = guide_legend(title.position="top", 
                               ncol = 2, nrow = 4, byrow = FALSE))
  
  print(array_prok[[one_strait]])
  
  }


```


### FIG Most abundant ASVs - Save as Grid
```{r}
   grid_fig_ASV <-  cowplot::plot_grid(array_prok[["Singapore"]],array_prok[["Johor"]], 
                                   array_euk[["Singapore"]],array_euk[["Johor"]], align="hv",
                                   # labels=c("A","B", "C", "D"),
                                   ncol=2, nrow=2, label_size = 20)


   ggsave("../fig/fig_top_asv_all.png", grid_fig_ASV, 
          height = 20, width = 20, dpi = 300)

```





## Bar plot of divisions per station


Note: some stations are completely missing heterotrophs (Only Opistokonta)

```{r bar_plot_division, fig.height=15, fig.width=10}

for (i in 1:3) {
  p <- plot_bar(ps_list$ps[[i]], x="sample_label", fill = "division") + 
       geom_bar(aes(color=division, fill=division), stat="identity", position="stack") + 
       ggtitle(str_c("Division level - ",ps_list$title[[i]]))  +
       theme(axis.text.y = element_text(size=10)) +
       theme(axis.text.x = element_text(angle=0, hjust = 0.5)) +
       coord_flip() +
      scale_fill_viridis(discrete=TRUE) + scale_color_viridis(discrete=TRUE)
  print(p)
}
```

## Bar plot of class per station

Only consider the abundant taxa

```{r bar_plot_class, fig.height=15, fig.width=10}

for (i in 1:3) {
  p <- plot_bar(ps_list_abundant$ps[[i]], x="sample_label", fill = "class") + 
       geom_bar(aes(color=class, fill=class), stat="identity", position="stack") + 
       ggtitle(str_c("Class level - ",ps_list_abundant$title[[i]])) +
       theme(axis.text.y = element_text(size=10))+
       theme(axis.text.x = element_text(angle=0, hjust = 0.5)) +
       coord_flip() +
      scale_fill_viridis(discrete=TRUE) + scale_color_viridis(discrete=TRUE)
  print(p)
}
```

## Compare by Straight, Site, Moonsoon (abundant OTUs only)

```{r bar_plot_factors, fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
for (factor in c("strait", "location", "monsoon")) {  
 for (i in 1:3) {
  ps_aggregate <- merge_samples(ps_list_abundant$ps[[i]], factor)
  ps_aggregate <- transform_sample_counts(ps_aggregate, function(x) 100*(x/sum(x)))
  p <- plot_bar(ps_aggregate, fill = "division") + 
      geom_col(aes(color=division, fill=division)) +
      ggtitle(str_c(ps_list_abundant$title[[i]], " - ", factor)) +
      theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
      ylab("%")+
      scale_fill_viridis(discrete=TRUE) + scale_color_viridis(discrete=TRUE)
  print(p)
 }
}
```

## Compare by Moonsoon in Singapore straight (abundant OTUs)

```{r bar_plot_monsoon, fig.height=4, fig.width=6, message=FALSE, warning=FALSE}

 for (i in 1:3) {
  ps_aggregate <- subset_samples(ps_list_abundant$ps[[i]], strait=="Singapore") 
  ps_aggregate <- merge_samples(ps_aggregate,"monsoon")
  ps_aggregate <- transform_sample_counts(ps_aggregate, function(x) 100*(x/sum(x)))
  p <- plot_bar(ps_aggregate, fill = "division") + 
      geom_col(aes(color=division, fill=division)) +
      ggtitle(str_c(ps_list_abundant$title[[i]], " - ", factor)) +
      theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
      ylab("%")+
      scale_fill_viridis(discrete=TRUE) + scale_color_viridis(discrete=TRUE)
  print(p)
 }
```


## Time series (abundant OTUs) aggregated for Singapore strait only

```{r bar_plot_date, fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
for (factor in c("date")) {  
 for (i in 1:3) {
  ps_aggregate <- subset_samples(ps_list_abundant$ps[[i]], strait=="Singapore")
  ps_aggregate <- merge_samples(ps_aggregate, "date")
  ps_aggregate <- transform_sample_counts(ps_aggregate, function(x) 100*(x/sum(x)))
  p <- plot_bar(ps_aggregate, fill = "division") + 
      geom_col(aes(color=division, fill=division)) +
      ggtitle(str_c(ps_list_abundant$title[[i]], " - ", factor)) +
      theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
      ylab("%")+
      scale_fill_viridis(discrete=TRUE) + scale_color_viridis(discrete=TRUE)
  print(p)
 }
}
```

## Time series (abundant OTUs) - Division level

```{r bar_plot_date_division, fig.height=6, fig.width=12, message=FALSE, warning=FALSE}
 for (i in 1:3) {
  ps_plot <- ps_list_abundant$ps[[i]]
  p <- plot_bar(ps_plot, x="date", fill = "division")+ 
      facet_grid(rows = vars(location)) + 
      geom_col(aes(color=division, fill=division)) +
      ggtitle(str_c(ps_list_abundant$title[[i]], " - Date")) +
      theme(axis.text.x = element_text(angle = 45, vjust=0.5, hjust=1)) +
      ylab("%") +
      xlim(as.POSIXct(as.Date(c("2017-01-01","2019-01-01" )))) + 
      geom_vline(xintercept = as.POSIXct(as.Date(c("2017-01-01","2018-01-01", "2019-01-01"))))+
      scale_fill_viridis(discrete=TRUE) + scale_color_viridis(discrete=TRUE)
  print(p)
 }
```


## Time series (abundant OTUs) - Genus level for Chlorophyta

```{r bar_plot_date_genus, fig.height=6, fig.width=12, message=FALSE, warning=FALSE}
 for (i in 3:3) {
  ps_plot <- subset_taxa(ps_list_abundant$ps[[i]], division == "Chlorophyta")
  p <- plot_bar(ps_plot, x="date", fill = "genus")+ 
      facet_grid(rows = vars(location)) + 
      geom_col(aes(color=genus, fill=genus)) +
      ggtitle(str_c(ps_list_abundant$title[[i]], " - Date")) +
      theme(axis.text.x = element_text(angle = 45, vjust=0.5, hjust=1)) +
      ylab("%") +
      xlim(as.POSIXct(as.Date(c("2017-01-01","2019-01-01" )))) + 
      geom_vline(xintercept = as.POSIXct(as.Date(c("2017-01-01","2018-01-01", "2019-01-01"))))+
      scale_fill_viridis(discrete=TRUE) + scale_color_viridis(discrete=TRUE)
  print(p)
 }
```


## Main species for each division (Eukaryotes - Autrotrophs)

```{r bar_plot_species, fig.height=8, fig.width=8}
p <- list()
for (one_division in c("Chlorophyta", "Dinoflagellata", "Ochrophyta", "Cryptophyta", "Haptophyta")){
  ps_subset <- subset_samples(ps_photo_abundant, strait != "Raffles Mari")
  ps_subset <- subset_taxa(ps_subset, division %in% one_division)
  p[[one_division]] <- plot_bar(ps_subset, x="species") + 
    facet_grid(rows = vars(strait), cols=vars(monsoon)) +
    geom_col() +
    theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
    ggtitle(str_c(one_division," - Abundant OTUs"))+
    coord_flip()
  print(p[[one_division]])  
}
    # grid_array <- list(p[["Chlorophyta"]],p[["Ochrophyta"]])
    # grid_layout <- rbind(c(NA,1,1,1,1,1),
    #                      c( 2,2,2,2,2,2))
    # grid_fig <- gridExtra::grid.arrange(grobs=grid_array, 
    #                                              layout_matrix=grid_layout, clip=FALSE, 
    #                                              padding = unit(0, "line"),
    #                                             as.table = FALSE) 
   grid_fig <-  cowplot::plot_grid(p[["Chlorophyta"]],p[["Ochrophyta"]], labels=c("A","B"), align="v", nrow=2)
   # ggsave("../fig/Fig_Species_Moonsoon.png", grid_fig, height = 10, width = 8, dpi = 300)
```

## Heatmaps


### Abundant OTUs

* Data are agglomarated at the genus level. Use function `tax_glom`

```{r heatmaps, fig.height=10, fig.width=10}

for (i in c(1)) {
  ps_heat <- tax_glom(ps_list_abundant$ps[[i]],taxrank = "family")
  p <- plot_heatmap(ps_heat, method = "NMDS", distance = "bray", 
               taxa.label = "family",  taxa.order="division",
               sample.label = "sample_label", sample.order="sample_label",
               low="beige", high="red", na.value="beige", 
               title=ps_list_abundant$title[[i]])
  print(p)
}
  

for (i in 2:3) {
  
  ps_heat <- tax_glom(ps_list_abundant$ps[[i]],taxrank = "genus")
  p <- plot_heatmap(ps_heat, method = "NMDS", distance = "bray", 
               taxa.label = "genus",  taxa.order="division",
               sample.label = "sample_label", sample.order="sample_label",
               low="beige", high="red", na.value="beige", 
               title=ps_list_abundant$title[[i]])
  print(p)
}


```

### Chlorophyta at species level
All ASVs considered (not only abundant)

```{r heatmap-chlorophyta, fig.height=10, fig.width=15}

  ps_heat <- subset_taxa(ps_photo, division == "Chlorophyta")
  
  ps_heat <- tax_glom(ps_heat,taxrank = "species")

  p <- plot_heatmap(ps_heat, method = "NMDS", distance = "bray", 
               taxa.label = "species",  taxa.order="species",
               sample.label = "sample_label", sample.order="sample_label",
               low="beige", high="red", na.value="beige",trans = scales::log10_trans(), 
               title="Mamiellophyceae in Singapore")
  print(p)
  

```

### Mamiello (Only Bathy, Ostreo and Micromonas) at genus level
All ASVs considered (not only abundant)

```{r heatmap_mamiello_1, fig.height=4, fig.width=10}

  ps_heat <- subset_taxa(ps_photo, genus %in% c("Ostreococcus", "Bathycoccus", "Micromonas"))
  
  ps_heat <- tax_glom(ps_heat,taxrank = "genus")

  p <- plot_heatmap(ps_heat, method = "NMDS", distance = "bray", 
               taxa.label = "genus",  taxa.order="genus",
               sample.label = "sample_label", sample.order="sample_label",
               low="beige", high="red", na.value="beige",trans = scales::log_trans(10), 
               title="Mamiellophyceae in Singapore")
  print(p)
  

```

### Mamiello (Only Bathy, Ostreo and Micromonas) at species level
All ASVs considered (not only abundant)

```{r heatmap_mamiello, fig.height=4, fig.width=10}

  ps_heat <- subset_taxa(ps_photo, genus %in% c("Ostreococcus", "Bathycoccus", "Micromonas"))
  
  ps_heat <- tax_glom(ps_heat,taxrank = "species")

  p <- plot_heatmap(ps_heat, method = "NMDS", distance = "bray", 
               taxa.label = "species",  taxa.order="species",
               sample.label = "sample_label", sample.order="sample_label",
               low="beige", high="red", na.value="beige",trans = scales::log_trans(10), 
               title="Mamiellophyceae in Singapore")
  print(p)
  

```

### Syndiniales at order level
All ASVs considered (not only abundant)

```{r heatmap_syndiniales_order, fig.height=4, fig.width=10}

  ps_heat <- subset_taxa(ps_euk, class %in% c("Syndiniales")) 
  ps_heat <- subset_samples(ps_heat, !is.na(Salinity))
  
  ps_heat <- tax_glom(ps_heat,taxrank = "order")

  p <- plot_heatmap(ps_heat, method = "NMDS", distance = "bray", 
               taxa.label = "order",  taxa.order="order",
               sample.label = "Salinity", sample.order="Salinity",
               low="beige", high="red", na.value="beige",trans = scales::log_trans(10), 
               title="Syndiniales in Singapore") +
      facet_grid(~location, scales="free_x")
  print(p)
  
  p <- plot_heatmap(ps_heat, method = "NMDS", distance = "bray", 
               taxa.label = "order",  taxa.order="order",
               sample.label = "Temperature", sample.order="Temperature",
               low="beige", high="red", na.value="beige",trans = scales::log_trans(10), 
               title="Syndiniales in Singapore") +
      facet_grid(~location, scales="free_x")
  print(p)
```

```{r heatmap_syndiniales_family, fig.height=10, fig.width=10}

  ps_heat <- subset_taxa(ps_euk, class %in% c("Syndiniales")) 
  ps_heat <- subset_samples(ps_heat, !is.na(Salinity))
  
  ps_heat <- tax_glom(ps_heat,taxrank = "family")

  p <- plot_heatmap(ps_heat, method = "NMDS", distance = "bray", 
               taxa.label = "family",  taxa.order="family",
               sample.label = "Salinity", sample.order="Salinity",
               low="beige", high="red", na.value="beige",trans = scales::log_trans(10), 
               title="Syndiniales in Singapore")+
      facet_grid(~location, scales="free_x")
  print(p)
  

```



### FIG - Most abundant OTUs in Johor vs Singapre and Prokaryotes vs Eukaryotes


```{r heatmaps_OTUs, fig.height=10, fig.width=14}

heat_prok <- list()
heat_euk <- list()

for (strait_one in c("Singapore", "Johor")) {
     
for (kingdom_one in list(c("Bacteria", "Archaea"), "Eukaryota")){

  otu_strait_one <- otu_strait %>% 
    filter(strait == strait_one & kingdom %in% kingdom_one) %>% 
    top_n(30, n_seq)

  ps_heat <- ps_all %>% 
    subset_samples(strait==strait_one) %>%
    subset_taxa(kingdom %in% kingdom_one)
  
  otu_table_keep <- data.frame(otu_table(ps_heat)) %>% 
    rownames_to_column("otu_id") %>% 
    filter(otu_id %in% otu_strait_one$otu_id) %>% 
    left_join(select(otu_strait_one, otu_id)) %>% 
    column_to_rownames("otu_id")
  
  tax_table_keep <- data.frame(tax_table(ps_heat)) %>% 
    rownames_to_column("otu_id") %>% 
    filter(otu_id %in% otu_strait_one$otu_id) %>% 
    left_join(select(otu_strait_one, otu_id, otu_label)) %>% 
    column_to_rownames("otu_id")

  OTU = otu_table(as.matrix(otu_table_keep), taxa_are_rows = TRUE)
  TAX = tax_table(as.matrix(tax_table_keep))
  samples = sample_data(ps_heat)
  
  ps_heat <- phyloseq(OTU, TAX, samples)    

  print(ps_heat)

  p <- plot_heatmap(ps_heat, method = "NMDS", distance = "bray", 
               taxa.label = "otu_label",  taxa.order="supergroup",
               sample.label = "sample_label", sample.order="sample_label",
               low="beige", high="red", na.value="beige",trans = scales::log10_trans(), 
               title=strait_one) +
    theme(axis.text.y = element_text(size = 16, angle = 0, hjust = 0, vjust = 0.5)) +
    theme(axis.text.x = element_text(size = 16)) +
    theme(axis.title = element_text(size = 0)) +
    theme(plot.title = element_text(size=22, hjust = 0.5)) +
    xlab("Sample") + ylab("")

  print(p)
  if (kingdom_one=="Eukaryota") heat_euk[[strait_one]] <- p 
  else heat_prok[[strait_one]] <- p

}}


   grid_fig_heat <-  cowplot::plot_grid(heat_prok[["Singapore"]],heat_prok[["Johor"]], align="hv",
                                   # labels=c("A","B"),
                                   ncol=1, nrow=2, label_size = 20)


   ggsave("../fig/fig_heat_prok.png", grid_fig_heat, 
          height = 20, width = 15, dpi = 300)

   grid_fig_heat <-  cowplot::plot_grid(heat_euk[["Singapore"]],heat_euk[["Johor"]], align="hv",
                                   # labels=c("A","B"),
                                   ncol=1, nrow=2, label_size = 20)


   ggsave("../fig/fig_heat_euk.png", grid_fig_heat, 
          height = 20, width = 15, dpi = 300)

```



## NMDS



Sample removed because they were pulling the NMDS
* PR2X16XS21 it has a single eukaryote (diatom bloom )
* RM13XS36 cause problem for bacteria
* PR11XS25 cause problem for hetero euks
* SBW11XS26 cause problem for hetero euks
* SBW13XS37 cause problem for hetero euks
* RM13XS36 cause problem for hetero euks


### Define function

* See comments inside functions (saved plots are different from displayed plots)

```{r}
ps_do_nmds <- function(ps_list) {

    plot_array <- list()
    
    for (i in 1:3) {
      ps_nmds <- ps_list$ps[[i]]
      
      # Remove samples with no reads
      ps_nmds <- prune_samples(sample_sums(ps_nmds)>0, ps_nmds)
      
      # Remove samples from Raffles and Johor
      ps_nmds <- subset_samples(ps_nmds, !(str_detect(strait, "Raffles Mari|Johor")))
      
      # Remove samples that caused problems (1= prok, 2=euk, 3=euk auto)
      if (i==1) {  ps_nmds <- prune_samples(!(sample_names(ps_nmds) %in% c("RM13XS36")), ps_nmds) } # Prokaryotes
      if (i %in% c(2,3)) {  ps_nmds <- prune_samples(!(sample_names(ps_nmds) %in% c("PR2X16SXS21")), ps_nmds)} # Eukaryotes
      if (i==4) {  ps_nmds <- prune_samples(!(sample_names(ps_nmds) %in% c("PR11XS25", "SBW11XS26", "SBW13XS37")), ps_nmds) } # Heterotrophs not used
       
      singa.ord <- ordinate(ps_nmds, "NMDS", "bray")
      
      # Fit environmental parameters
       env_var <- sample_variables(ps_nmds)
       env_matrix <- get_variable(ps_nmds,c("Chl", "Temperature", "Salinity", "Phosphate", "Silicate", "DIN", "BAC") )
       env_fit <- vegan::envfit(singa.ord, env = env_matrix, perm = 999, na.rm=TRUE)
       env_arrows <- data.frame(env_fit$vectors$arrows*sqrt(env_fit$vectors$r)) %>% rownames_to_column(var="parameter")
      
      
      nmds_samples <- data.frame(singa.ord[["points"]], 
                                 get_variable(ps_nmds,c("monsoon", "strait", "location", "location_label") ) ) %>%
                      rownames_to_column(var="sample") 

      # Factor to move the labels
      nudge_x <- max(nmds_samples$MDS1)*0.08
      nudge_y <- max(nmds_samples$MDS2)*0.08
      xy_max = max(c(nmds_samples$MDS1, nmds_samples$MDS2))*1.5
      xy_min = min(c(nmds_samples$MDS1, nmds_samples$MDS2))*1.5
      factor <- 3 # for vectors for euks
      if (i==1) {factor <- 1.5} # for vectors for proks
      print(factor)
      
        p <- plot_ordination(ps_nmds, singa.ord, type="samples", color="monsoon",
                      shape="strait", title=ps_list$title[[i]]) +
           geom_point(aes(shape=strait, color=monsoon), size=3.5)  +
           scale_color_viridis(discrete = TRUE) +
           scale_shape_manual(values = c(15,16)) +
           geom_text(aes(label=location_label, color=monsoon),
                     nudge_x=nudge_x, nudge_y=nudge_y,
                     check_overlap = TRUE, size=2) +
           theme_bw() +
           geom_segment(data=env_arrows, aes(x=0,xend=NMDS1*factor, y=0,yend=NMDS2*factor), inherit.aes = FALSE,
                        arrow = arrow(length = unit(0.5, "cm")), colour = "black")  +
           geom_text(data=env_arrows, aes(x=NMDS1*factor, y=NMDS2*factor, label=parameter), inherit.aes = FALSE ,
                     hjust=-0.2, vjust=-0.2, size=3) 
      print(singa.ord)
      plot_array[[i]] <- p
      print(p)
      
      # The following lines can be used if you want to avoid using the pjhyloseq functions to plot the data.
      # Notes : - must use inherit.aes = FALSE to add some extra layers
      #         - the saved plots have a different scale for the added layer than the displaued plot can figure out 
           # ggplot()+
           # coord_fixed() + 
           # xlim(xy_min, xy_max) + ylim(xy_min, xy_max) +
           # geom_point(data=nmds_samples, aes(x=MDS1, y=MDS2, shape=strait, color=monsoon), size=5)  +
           # geom_text(data=nmds_samples, aes(x=MDS1, y=MDS2, label=location_label, color=monsoon),
           #           nudge_x=nudge_x, nudge_y=nudge_y,
           #           check_overlap = FALSE, size=2) +           
           # ggtitle(ps_list$title[[i]]) +
      
       p <- plot_ordination(ps_nmds, singa.ord, type="taxa", color="division", title=ps_list$title[[i]]) +
       scale_color_viridis(discrete = TRUE) +
       geom_point(size=3) +
       theme_bw() +
       geom_segment(data=env_arrows, aes(x=0,xend=NMDS1*factor, y=0,yend=NMDS2*factor), inherit.aes = FALSE,
                    arrow = arrow(length = unit(0.5, "cm")), colour = "black")  +
       geom_text(data=env_arrows, aes(x=NMDS1*factor, y=NMDS2*factor, label=parameter), inherit.aes = FALSE ,
                 hjust=-0.2, vjust=-0.2, size=3) 
       
       print(p)
       plot_array[[i+3]] <- p
    }
  return(plot_array)
  }
```

### All OTUs 

```{r nmds_all, fig.height=8, fig.width=8}

  plot_array_all <- ps_do_nmds(ps_list)  
  
```

### Abundant OTUs
```{r  nmds_abundant, fig.height=8, fig.width=8}

  plot_array_abundant <- ps_do_nmds(ps_list_abundant)   

```
### MSDS graph

```{r}
    grid_fig <- cowplot::plot_grid(plot_array_abundant[[1]], plot_array_abundant[[4]], 
                                   plot_array_abundant[[3]], plot_array_abundant[[6]], ncol=2,
                                   labels = c("A", "B", "C", "D"),
                                   align="hv", label_size = 20)
   grid_fig
   # ggsave("../fig/Fig_NMDS-Singapore.png", grid_fig, height = 12, width = 18, dpi = 300)
```



## Network analysis

```{r network, fig.height=6, fig.width=10}

for (i in 1:3) {
  
  ps_nmds <- ps_list_abundant$ps[[i]]
  
  # Remove samples with no reads
  ps_nmds <- prune_samples(sample_sums(ps_nmds)>0, ps_nmds)
  
  # Remove samples that caused problems (1= prok, 2=euk, 3=euk auto)
  if (i==1) {  ps_nmds <- prune_samples(!(sample_names(ps_nmds) %in% c("RM13XS36")), ps_nmds) } # Prokaryotes
  if (i %in% c(2,3)) {  ps_nmds <- prune_samples(!(sample_names(ps_nmds) %in% c("PR2X16SXS21")), ps_nmds)} # Eukaryotes
  if (i==4) {  ps_nmds <- prune_samples(!(sample_names(ps_nmds) %in% c("PR11XS25", "SBW11XS26", "SBW13XS37")), ps_nmds) } # Heterotrophs not used

  if (i > 1) {
  p <- plot_net(ps_nmds, distance = "(A+B-2*J)/(A+B)", type = "taxa", 
           maxdist = 0.4, color="class", point_label="genus") +
       ggtitle(ps_list_abundant$title[[i]])
  }

  else {
  p <- plot_net(ps_nmds, distance = "(A+B-2*J)/(A+B)", type = "taxa", 
           maxdist = 0.4, color="class", point_label="family") +
       ggtitle(ps_list_abundant$title[[i]])
  }

  print(p)
}

```


## Differential expression (DESeq2)

Love, M.I., Huber, W. & Anders, S. 2014. Moderated estimation of fold change and  dispersion for RNA-seq data with DESeq2. Genome Biol. 15:550.

* Follow: 
    * https://joey711.github.io/phyloseq-extensions/DESeq2.html
    * https://bioconductor.org/packages/devel/bioc/vignettes/phyloseq/inst/doc/phyloseq-mixture-models.html

* Start from **raw counts** (not normalized) : `ps_all_raw `

```{r}
library("DESeq2")
packageVersion("DESeq2")
```



Block 1 is need to avoid the following error
> Error in estimateSizeFactorsForMatrix(counts(object), locfunc = locfunc, : every gene contains at least one zero, cannot compute log geometric means

See: https://github.com/joey711/phyloseq/issues/387


```{r, message=FALSE}
ps <- ps_all_raw 
ps <- subset_samples(ps, strait=="Singapore" & monsoon %in% c("NE", "SW"))

ps

diagdds = phyloseq_to_deseq2(ps, ~ monsoon)

# ps <- subset_samples(ps, strait %in% c("Singapore","Johor"))
# diagdds = phyloseq_to_deseq2(ps, ~ strait)

# Block 1 Start
# Estimate geometric mean with zeros

gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}
geoMeans = apply(counts(diagdds), 1, gm_mean)
diagdds = estimateSizeFactors(diagdds, geoMeans = geoMeans)

# Block 1 end

diagdds = DESeq(diagdds, fitType="local")
```



```{r}
res = results(diagdds, cooksCutoff = FALSE)
alpha = 0.01
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(ps)[rownames(sigtab), ], "matrix"))
head(sigtab)
```

### FIG 

```{r fig.height=14, fig.width=14}

sigtab_plot <- sigtab %>% 
  rownames_to_column(var = "otu_id") %>% 
  top_n(-90, wt=otu_id) %>% 
  mutate(label= case_when(
         kingdom == "Eukaryota" ~ str_c(otu_id, class,species, sep="_"),
         TRUE ~ str_c(otu_id, class,genus, sep="_")),
         supergroup_label = str_c(str_sub(kingdom, 1,1), supergroup, sep="_")) %>% 
  arrange(desc(otu_id)) %>% 
  rowid_to_column()




g <- ggplot(sigtab_plot, aes(x = reorder(label, log2FoldChange), y = log2FoldChange, color = kingdom, alpha=log(rowid))) +
  theme_dviz_grid() + 
  geom_point(size = 6) +
  theme(axis.text.x = element_text(size = 16, angle = 0, hjust = 0.5, vjust = 0.5 ),
        axis.text.y = element_text(size = 16, angle = 0, hjust = 0, vjust = 0.5 ),
        legend.title = element_text(size = 24),
        legend.text = element_text(size = 20)) +
  guides(alpha = FALSE) +
  ylab("log2 fold change - NE <--> SW") + xlab("ASV") +
  scale_color_viridis_d(option = "C", name = "Kingdom") +
  coord_flip()

g

 ggsave(filename="../fig/fig_DESeq2.png", g, 
        height = 20, width = 20, dpi = 300)

```

### Do a histogram plot for all ASV significantly different

```{r fig.height=6, fig.width=6}
  p <- list()
  for (asv_one in sigtab_plot$otu_id){
  ps_one <- prune_taxa(asv_one, ps)
  p[[asv_one]] <- plot_bar(ps_one, x="sample_label", fill = "class") +
       geom_bar(aes(color=class, fill=class), stat="identity", position="stack") +
       ggtitle(asv_one)  +
       theme(axis.text.y = element_text(size=10)) +
       theme(axis.text.x = element_text(size=8, angle=0, hjust = 0.5)) +
       coord_flip() +
       scale_fill_viridis(discrete=TRUE) + scale_color_viridis(discrete=TRUE)
  print(p[[asv_one]])
  }

  p_grid <- gridExtra::grid.arrange(grobs=p, nrow=18, ncol=4)
  ggsave(filename="../fig/fig_DESeq2_grid.png", p_grid, 
        height = 45, width = 20, dpi = 400)

```

